```{r}
library(glmnet)
library(ggplot2)
library(WVPlots)
```

Modeling functions
```{r}
# assuming xframe is entirely numeric
ridge_predict_function = function(model) {
  # to get around the 'unfullfilled promise' leak. blech.
  force(model) 
  function(xframe) {
    as.numeric(predict(model, newx=as.matrix(xframe), type="response"))
  }
}

# assuming the xframe is entirely numeric
# if there are categories, we would have to use 
#   model_matrix, or something
# assuming family is one of c("binomial", "gaussian")
#   should have a check for that
ridge_model = function(xframe, y, family="binomial") {
  model = glmnet(as.matrix(xframe), y, alpha=0, 
                 lambda=0.001, family=family)
  list(coef = coef(model),
       deviance = deviance(model),
       predfun = ridge_predict_function(model)
  )
}
```

```{r model}
# loads vars (names of vars), yVar (name of y column), 
# dTrainS, dTestS
load("wideData.rData")

model = ridge_model(dTrainS[,vars], dTrainS[[yVar]])

names = rownames(model$coef)
coefs = as.vector(model$coef)
names(coefs) = names
ord = order(abs(coefs), decreasing=TRUE)
coefs[ord]

testpred = model$predfun(dTestS[,vars])
dTestS$pred = testpred

DoubleDensityPlot(dTestS, "pred", yVar, 
                  title="Model Score Distribution")
ROCPlot(dTestS, "pred", yVar, title="ROC plot")
GainCurvePlot(dTestS, "pred", yVar, title="Gain plot")

```

Evaluate the positive examples. Do we detect churn in time?

*Note: from the implementation, we know that things in "bad" state
will leave in about 5 days. things not in bad state will leave in about 15 days.*

```{r evaluate}
dTestS$predictedToLeave = dTestS$pred>0.5

# confusion matrix
cmat = table(pred=dTestS$predictedToLeave, actual=dTestS[[yVar]])
cmat
recall = cmat[2,2]/sum(cmat[,2])
recall
precision = cmat[2,2]/sum(cmat[2,])
precision

# make daysToX finite. The idea is that the live-forevers should be rare
isinf = dTestS$daysToX==Inf
maxval = max(dTestS$daysToX[!isinf])
dTestS$daysToX = with(dTestS, ifelse(daysToX==Inf, maxval, daysToX))

# how long on average until flagged customers leave?
posmean = mean(dTestS[dTestS$predictedToLeave, "daysToX"])
posmean

# how long on average until unflagged customers leave?
negmean = mean(dTestS[!dTestS$predictedToLeave, "daysToX"])
negmean

print(negmean-posmean) # theoretically, the difference can't be above 10
```

```{r drilldown}
# xvar is integral
dist_and_mean = function(frm, xvar, title, meanlabel) {
  meanval = mean(frm[[xvar]])
  mode = max(table(frm[[xvar]]))
  print(paste("Mean days until exit:", meanval))
  
  DiscreteDistribution(frm, xvar, title=title) +
    geom_vline(xintercept=meanval, color="blue", linetype=2)+
    annotate("text", x=meanval, y=mode,
             hjust=0, vjust=0,
             label=meanlabel,
             color="blue")
}


# how long until true positives (customers flagged as leaving who really do leave) leave?
tpfilter = dTestS$predictedToLeave & dTestS[[yVar]]
dist_and_mean(dTestS[tpfilter,], "daysToX", 
              "Distribution of days til exit, true positives",
              "mean days til exit") 
              

# same plot as above for all positives and all negatives
dist_and_mean(dTestS[dTestS$predictedToLeave,], "daysToX",
               "Distribution of days til exit, predicted positive",
              "mean days til exit") + geom_vline(xintercept=7, color="red")

dist_and_mean(dTestS[!dTestS$predictedToLeave,], "daysToX",
               "Distribution of days til exit, predicted negative",
              "mean days til exit") + geom_vline(xintercept=7, color="red")

# how soon do the actual positives that we missed leave?
falsenegmean = mean(dTestS[!dTestS$predictedToLeave & dTestS[[yVar]], "daysToX"])
falsenegmean

# alternative to the double density plot above;
# plots days to churn, instead of model score
ggplot(dTestS, aes_string(x="daysToX", color="predictedToLeave")) +
  geom_density(adjust=0.5) + 
  geom_vline(xintercept=7, color="darkgray") +
  geom_vline(xintercept=posmean, color="cyan4", linetype=2) +
  geom_vline(xintercept=negmean, color="red", linetype=2) + 
  ggtitle("Days to Exit")

ScatterBoxPlot(dTestS, "predictedToLeave", "daysToX", pt_alpha=0.2,
               title="Distribution of days til exit")



